---
title: "LEGO sets"
author: "Pedro Jordano"
date: "febrero 18, 2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### LEGO sets diversity and connectivity analysis

This is an analysis of LEGO sets patterns of component parts: their diversity and connectivity forms. 

```{r packages, message=FALSE}
library(dplyr)
library(igraph)
library(gt)
library(stringr)
library(vegan)
library(here)
library("png")                      # Load png package
# install.packages("patchwork")       # Install patchwork package
library("patchwork")                # Load patchwork

```

## Input data

Input inventory datasets from [BrickLink][http://www.bricklink.com]

```{r input_lego }
# Input from local drive; previously downloaded files from Rebrickable.
colors <- readr::read_csv(here::here("data/extdata/colors.csv"))
inventories <- readr::read_csv("data/extdata/inventories.csv")
inventory_parts <- readr::read_csv("data/extdata/inventory_parts.csv")
inventory_sets <- readr::read_csv("data/extdata/inventory_sets.csv")
part_categories <- readr::read_csv("data/extdata/part_categories.csv")
parts <- readr::read_csv("data/extdata/parts.csv")
sets <- readr::read_csv("data/extdata/sets.csv")
themes <- readr::read_csv("data/extdata/themes.csv")
elements <- readr::read_csv("data/extdata/elements.csv")
inventory_minifigs <- readr::read_csv("data/extdata/inventory_minifigs.csv")
minifigs <- readr::read_csv("data/extdata/minifigs.csv")
part_relationships <- readr::read_csv("data/extdata/part_relationships.csv")

# Direct download from Rebrickable.
urlthemes <- "https://cdn.rebrickable.com/media/downloads/themes.csv.gz"
urlcolors <- "https://cdn.rebrickable.com/media/downloads/colors.csv.gz"
urlpart_categories<- "https://cdn.rebrickable.com/media/downloads/part_categories.csv.gz"
urlparts<- "https://cdn.rebrickable.com/media/downloads/parts.csv.gz"
urlsets<- "https://cdn.rebrickable.com/media/downloads/sets.csv.gz?1678349282.4992967"
urlpart_relationships<- "https://cdn.rebrickable.com/media/downloads/part_relationships.csv.gz"
urlelements<- "https://cdn.rebrickable.com/media/downloads/elements.csv.gz"
urlminifigs<- "https://cdn.rebrickable.com/media/downloads/minifigs.csv.gz"
urlinventories <- "https://cdn.rebrickable.com/media/downloads/inventories.csv.gz"
urlinventory_parts<- "https://cdn.rebrickable.com/media/downloads/inventory_parts.csv.gz"
urlinventory_sets<- "https://cdn.rebrickable.com/media/downloads/inventory_sets.csv.gz"
urlinventory_minifigs<- "https://cdn.rebrickable.com/media/downloads/inventory_minifigs.csv.gz"

##----------------------------
dwld <- function(url) {
    tmp <- tempfile()
    download.file(url,tmp)  
    read.csv(
    gzfile(tmp),
    sep=",",
    header=TRUE,
    stringsAsFactors=T)
      #names(data)[1] <- sub("X\\.","",names(data)[1])
}
#-----------------------------
themes<- dwld(urlthemes)
colors <- dwld(urlcolors)
inventories <- dwld(urlinventories)
inventory_parts <- dwld(urlinventory_parts)
inventory_sets <- dwld(urlinventory_sets)
part_categories <- dwld(urlpart_categories)
parts <- dwld(urlparts)
sets <- dwld(urlsets)
themes <- dwld(urlthemes)
elements <- dwld(urlelements)
inventory_minifigs <- dwld(urlinventory_minifigs)
minifigs <- dwld(urlminifigs)
part_relationships <- dwld(urlpart_relationships)

```


Create joined datasets with different informatioon on sets and parts.
```{r joined_datasets}
# Joins:
# 0 inventory_parts
# 1 . , inventories
# 2 . , parts
# 3 . , part_categories
# 4 . , colors
# 5 . , sets
# 6 . , themes

library("dplyr")
lego_parts <- left_join(inventory_parts, inventories,
  by=c("inventory_id" = "id")) %>%
  left_join(parts, by="part_num") %>%
  left_join(part_categories, by=c("part_cat_id"="id")) %>%
  rename("part_name"=name.x, "part_category"=name.y) %>%
  left_join(colors, by=c("color_id"="id")) %>%
  rename(color_name="name") %>%
  left_join(sets, by="set_num") %>%
  left_join(themes, by=c("theme_id" = "id")) %>%
  rename("set_name"=name.x, "theme_name"=name.y)

lego_parts <- rename(lego_parts, "set_img_url"=img_url.y, "part_img_url"=img_url.x)

lego_parts<- lego_parts %>% select(theme_name, 
                      theme_id, 
                      set_img_url,
                      set_num, 
                      set_name, 
                      year, 
                      num_parts,
                      inventory_id, 
                      part_cat_id,
                      part_category,
                      part_num, 
                      part_name,
                      quantity,
                      part_img_url,
                      part_material,
                      color_id,
                      color_name,
                      is_trans,
                      rgb,
                      parent_id,
                      is_spare,
                      version)
write_csv(lego_parts, "data/extdata/lego_parts.csv")

lego_sets <- left_join(sets, themes, by=c("theme_id" = "id"))
lego_sets <- rename(lego_sets, "set_name"=name.x, "theme_name"=name.y,
                    "set_img_url"= img_url)
write_csv(lego_sets, "data/extdata/lego_sets.csv")

# main<- left_join(dleft06, elements, by=c("part_num","color_id"))
# names(main)
# dim(main)
# write_csv(main,"data/extdata/main.csv")

```

## Reationships between no. parts in set and no. part types
```{r n_parts-vs-n_part-types }

by_set_part<- Starwars %>% 
    dplyr::group_by(set_name, part_name)
tt<- by_set_part %>% 
    dplyr::count(part_name) 
by_set_nparts_ntypes<- tt %>% 
    group_by(set_name) %>% 
    dplyr::summarize(nparts= sum(n))

ttt<-  tt %>% 
    group_by(set_name) %>% 
    tally()
colnames(ttt)<- c("set_name","ntypes")
by_set_nparts_ntypes<- as.data.frame(cbind(by_set_nparts_ntypes,ttt[,2]))

by_set_nparts_ntypes<- by_set_nparts_ntypes %>% 
         dplyr::filter(nparts>10)

# Plots of no. total part in set vs no. part types.
# 
plot(by_set_nparts_ntypes$nparts,by_set_nparts_ntypes$ntypes)
gg <- ggplot(by_set_nparts_ntypes, aes(x = nparts, y = ntypes)) +
    geom_point()                           +
    scale_x_log10() + scale_y_log10()      +
    labs(x = "Number of LEGO parts", 
         y = "Number of LEGO part types")  +
         ggtitle("Starwars")               +
    theme_bw()
gg
# readPNG function to load an image that we have stored on our computer:
my_image <- readPNG("images/sw_logo.png", native = TRUE)

# Use the inset_element function to add our picture to our ggplot2 graphic:
ggp_image <- gg +                  # Combine plot & image
  inset_element(p = my_image,
                left = 0.15,
                bottom = 0.75,
                right = 0.4,
                top = 0.9, align_to = 'full', clip=T)
ggp_image                           # Draw combined plot


```

## Diversity

```{r subsets}
# Anakin Skywalker and Sebulba's Podracers
AnaSebPodRac_7962 <- lego_parts %>%
    dplyr::filter(set_num == "7962-1") 

tt<-AnaSebPodRac_7962  %>% 
    group_by(part_num)  %>% 
    dplyr::summarize(n = n())

##################
# SUBSETS
# 
lego_parts %>%
    dplyr::filter(theme_name == "Star Wars") 

podracers<- lego_parts %>%
         dplyr::filter(grepl('odracer', set_name))
microfighter<- lego_parts %>%
         dplyr::filter(grepl('microfighter', set_name))

Starwars<- lego_parts %>%
         dplyr::select(set_num, set_name, year, theme_name, theme_id, inventory_id, 
                       part_num, part_name, quantity, num_parts)  %>% 
         dplyr::filter(theme_name == "Star Wars")

Creator<- lego_parts %>%
        dplyr::select(set_num, set_name, year, theme_name, theme_id, inventory_id, 
                       part_num, part_name, quantity, num_parts) %>% 
        dplyr::filter(grepl("Creator", theme_name))

```

```{r diversity_themes}
# Groups of models
stw_H<- Starwars %>%
    group_by(set_name) %>%
    dplyr::summarize(H= vegan::diversity(table(quantity), 
                 index = "shannon", MARGIN = 1, base = exp(1)))
stw_Simp<- Starwars %>%
    group_by(set_name) %>%
    dplyr::summarize(simp= vegan::diversity(table(quantity), "simpson")) 
stw_invSimp<- Starwars %>%
    group_by(set_name) %>%
    dplyr::summarize(invSimp= vegan::diversity(table(quantity), 
                 index = "inv")) 
stw_unbSimp<- Starwars %>%
    group_by(set_name) %>%
    dplyr::summarize(unbias.simp= vegan::rarefy(table(quantity), 
                     2) - 1) 
# stw_alpha<- Starwars %>%
#     group_by(set_name) %>%
#     dplyr::summarize(alpha= vegan::fisher.alpha(table(quantity))) 
stw_S<- Starwars %>%
    group_by(set_name) %>%
    dplyr::summarize(S= vegan::specnumber(table(quantity)))
stw_J<- as.data.frame(cbind(set_name=stw_H$set_name, J= stw_H$H/log(stw_S$S)))
#(J= H/log(S))


#-----------------------------------
Starwars_div<- as.data.frame(cbind(
                            stw_H, 
                            simp= stw_Simp$simp,
                            invSimp= stw_invSimp$invSimp,
                            unbias.simp= stw_unbSimp$unbias.simp,
                            S= stw_S$S,
                            J= stw_J$J
                            ))

tt<- as.data.frame(table(Starwars$set_name))
tt <- rename(tt, "set_name"=Var1, "num_parts"=Freq)
tt<- inner_join(tt, Starwars[,1:8], by = "set_name",multiple="first")
Starwars_div<-  as.data.frame(cbind(tt,Starwars_div[,-1]))

Starwars_div<- Starwars_div %>% # Filtering out H'=0 records.
        dplyr::filter(H != 0)

# Plots of H' vs no. parts for sets.
# 
plot(Starwars_div$num_parts,Starwars_div$H)
gg <- ggplot(Starwars_div, aes(x = num_parts+1, y = H+1)) +
    geom_point() +
    scale_x_log10() + scale_y_log10() +
    theme_bw()
gg

#------------------------------------------------------------------------

```

#### Diversity for specific sets

```{r diversity_set, echo=FALSE}
# For a specific model.
H <- vegan::diversity(table(AnaSebPodRac_7962$part_num))

H <- vegan::diversity(table(AnaSebPodRac_7962$part_num), 
                 index = "shannon", MARGIN = 1, base = exp(1))

simp <- vegan::diversity(table(AnaSebPodRac_7962$part_num), "simpson")
invsimp <- vegan::diversity(table(AnaSebPodRac_7962$part_num), "inv")
## Unbiased Simpson (Hurlbert 1971, eq. 5) with rarefy:
unbias.simp <- vegan::rarefy(table(AnaSebPodRac_7962$part_num), 2) - 1

## Fisher alpha
alpha <- vegan::fisher.alpha(table(AnaSebPodRac_7962$part_num))
S <- vegan::specnumber(table(AnaSebPodRac_7962$part_num)) ## rowSums(Qty > 0) does the same...
J <- H/log(S)

## Plot all
cbind(H, simp, invsimp, unbias.simp, alpha, S, J)

```

### Diversity analysis for all LEGO sets.

```{r diversity_all_sets}
# Sets and Themes.
# Subsetting main themes (those with ≥ 10 sets).
names(lego_sets)
subset_themes_n10<- lego_sets %>% # Subsetting main themes (those with ≥ 10 sets).
                                  # There are 51 themes with at least 10 parts,
                                  # and are not Minifigures, Value Packs, or
                                  # Audio and Visual Media.
     dplyr::select(theme_name, theme_id, set_name, year, num_parts) %>% 
     dplyr::group_by(theme_name) %>%
     dplyr::summarise(sum_parts = sum(num_parts), n = n()) %>% 
     dplyr::filter(sum_parts> 0) %>% 
     dplyr::filter(!grepl('Audio and', theme_name)) %>% 
     dplyr::filter(!grepl('Minifig', theme_name)) %>% 
     dplyr::filter(!grepl('Key Chain', theme_name)) %>% 
     dplyr::filter(!grepl('Books', theme_name)) %>% 
     dplyr::filter(!grepl('Service Packs', theme_name)) %>% 
     dplyr::filter(!grepl('Stationery and Office Supplies', theme_name)) %>% 
     dplyr::filter(!grepl('Duplo and Explore', theme_name)) %>% 
     dplyr::filter(!grepl('Houseware', theme_name)) %>% 
     dplyr::filter(!grepl('Clocks and Watches', theme_name)) %>% 
     dplyr::filter(!grepl('Storage', theme_name)) %>% 
     dplyr::filter(!grepl('Bulk Bricks', theme_name)) %>% 
     dplyr::filter(!grepl('Duplo', theme_name)) %>% 
     dplyr::filter(!grepl('Magnets', theme_name)) %>% 
     dplyr::filter(!grepl('Primo', theme_name)) %>% 
     dplyr::filter(!grepl('Tabletop Games and Puzzles', theme_name)) %>% 
     dplyr::filter(!grepl('LEGO Originals', theme_name)) %>% 
     dplyr::filter(!grepl('Power Functions', theme_name)) %>% 
     dplyr::filter(!grepl('Powered Up', theme_name)) %>% 
     dplyr::filter(!grepl('Books', theme_name)) %>% 
     dplyr::filter(!grepl('Books', theme_name)) %>% 
     dplyr::filter(!grepl('Value', theme_name)) %>% 
     dplyr::filter(n > 9) %>% dplyr::arrange(desc(n))
print(subset_themes_n10, n=300)

# Select rows in lego_parts that match those sets listed in 'subset_themes_n10'.
#
subset_parts_themes_n10<- lego_parts %>% 
     # Subsetting main themes (those with ≥ 10 sets).
     # Here I filter those themes not included in 
     # subset_themes_n10, but using the whole list
     # of parts.
     select(theme_name, 
                      theme_id, 
                      set_img_url,
                      set_num, 
                      set_name, 
                      year, 
                      num_parts,
                      inventory_id, 
                      part_cat_id,
                      part_category,
                      part_num, 
                      part_name,
                      quantity,
                      part_img_url,
                      part_material,
                      color_id,
                      color_name,
                      is_trans,
                      rgb,
                      parent_id,
                      is_spare,
                      version) %>% 
     #dplyr::group_by(theme_name, set_name) %>%
     #dplyr::summarise(sum_parts = sum(num_parts), n = n()) %>% 
     #dplyr::filter(sum_parts> 0) %>% 
     dplyr::filter(!grepl('Audio and', theme_name)) %>% 
     dplyr::filter(!grepl('Minifig', theme_name)) %>% 
     dplyr::filter(!grepl('Key Chain', theme_name)) %>% 
     dplyr::filter(!grepl('Books', theme_name)) %>% 
     dplyr::filter(!grepl('Service Packs', theme_name)) %>% 
     dplyr::filter(!grepl('Stationery and Office Supplies', theme_name)) %>% 
     dplyr::filter(!grepl('Duplo and Explore', theme_name)) %>% 
     dplyr::filter(!grepl('Houseware', theme_name)) %>% 
     dplyr::filter(!grepl('Clocks and Watches', theme_name)) %>% 
     dplyr::filter(!grepl('Storage', theme_name)) %>% 
     dplyr::filter(!grepl('Bulk Bricks', theme_name)) %>% 
     dplyr::filter(!grepl('Duplo', theme_name)) %>% 
     dplyr::filter(!grepl('Magnets', theme_name)) %>% 
     dplyr::filter(!grepl('Primo', theme_name)) %>% 
     dplyr::filter(!grepl('Tabletop Games and Puzzles', theme_name)) %>% 
     dplyr::filter(!grepl('LEGO Originals', theme_name)) %>% 
     dplyr::filter(!grepl('Power Functions', theme_name)) %>% 
     dplyr::filter(!grepl('Powered Up', theme_name)) %>% 
     dplyr::filter(!grepl('Books', theme_name)) %>% 
     dplyr::filter(!grepl('Books', theme_name)) %>% 
     dplyr::filter(!grepl('Value', theme_name)) #%>% 
     #dplyr::filter(n > 9) %>% dplyr::arrange(desc(n))

#------------------------------------------------------------------------------
# Groups of models
# ALL SETS.
all_H<- lego_parts %>%
    group_by(set_name) %>%
    dplyr::summarize(H= vegan::diversity(table(quantity), 
                 index = "shannon", MARGIN = 1, base = exp(1)))
all_Simp<- lego_parts %>%
    group_by(set_name) %>%
    dplyr::summarize(simp= vegan::diversity(table(quantity), "simpson")) 
all_invSimp<- lego_parts %>%
    group_by(set_name) %>%
    dplyr::summarize(invSimp= vegan::diversity(table(quantity), 
                 index = "inv")) 
all_unbSimp<- lego_parts %>%
    group_by(set_name) %>%
    dplyr::summarize(unbias.simp= vegan::rarefy(table(quantity), 
                     2) - 1) 
# stw_alpha<- Starwars %>%
#     group_by(set_name) %>%
#     dplyr::summarize(alpha= vegan::fisher.alpha(table(quantity))) 
all_S<- lego_parts %>%
    group_by(set_name) %>%
    dplyr::summarize(S= vegan::specnumber(table(quantity)))
all_J<- as.data.frame(cbind(set_name=all_H$set_name, J= all_H$H/log(all_S$S)))
#(J= H/log(S))

#------------------------------------------------------------------------------
# ALL SETS INCLUDED. 
allsets_div<- as.data.frame(cbind(
                            all_H, 
                            simp= all_Simp$simp,
                            invSimp= all_invSimp$invSimp,
                            unbias.simp= all_unbSimp$unbias.simp,
                            S= all_S$S,
                            J= all_J$J
                            ))

tt<- as.data.frame(table(lego_parts$set_name))
tt <- rename(tt, "set_name"=Var1, "num_parts"=Freq)
tt<- inner_join(tt, lego_parts[,c(4,5,6,1,2,8,11,12)], by = "set_name",multiple="first")

# allsets_div<-  as.data.frame(cbind(tt,allsets_div))
allsets_div <- left_join(tt, allsets_div,
  by="set_name") 

allsets_div<- allsets_div %>% # Filtering out H'=0 records.
        dplyr::filter(H > 0.5)  %>% 
        dplyr::filter(num_parts> 15)

# Plots of H' vs no. parts for sets.
# 
plot(allsets_div$num_parts,allsets_div$H)
gg <- ggplot(allsets_div, aes(x = num_parts, y = H)) +
    geom_point() +
    scale_x_log10() + scale_y_log10() +
    theme_bw()
gg
#------------------------------------------------------------------------------

```

### Analyses for part-related themes; subsetted

Here I subsetted those themes that have actaul part-related sets., i.e., 
excluding themes like Books, Gear, etc.

I'm usismg the whole part list in order to estimate diversity parameters on sets.

```{r}
# Select rows in lego_parts that match those sets listed in 'subset_themes_n10'.
#
subset_parts_themes_n10<- lego_parts %>% 
     # Subsetting main themes (those with ≥ 10 sets).
     # Here I filter those themes not included in 
     # subset_themes_n10, but using the whole list
     # of parts.
     select(theme_name, 
                      theme_id, 
                      set_img_url,
                      set_num, 
                      set_name, 
                      year, 
                      num_parts,
                      inventory_id, 
                      part_cat_id,
                      part_category,
                      part_num, 
                      part_name,
                      quantity,
                      part_img_url,
                      part_material,
                      color_id,
                      color_name,
                      is_trans,
                      rgb,
                      parent_id,
                      is_spare,
                      version) %>% 
     #dplyr::group_by(theme_name, set_name) %>%
     #dplyr::summarise(sum_parts = sum(num_parts), n = n()) %>% 
     #dplyr::filter(sum_parts> 0) %>% 
     dplyr::filter(!grepl('Audio and', theme_name)) %>% 
     dplyr::filter(!grepl('Minifig', theme_name)) %>% 
     dplyr::filter(!grepl('Key Chain', theme_name)) %>% 
     dplyr::filter(!grepl('Books', theme_name)) %>% 
     dplyr::filter(!grepl('Service Packs', theme_name)) %>% 
     dplyr::filter(!grepl('Stationery and Office Supplies', theme_name)) %>% 
     dplyr::filter(!grepl('Duplo and Explore', theme_name)) %>% 
     dplyr::filter(!grepl('Houseware', theme_name)) %>% 
     dplyr::filter(!grepl('Clocks and Watches', theme_name)) %>% 
     dplyr::filter(!grepl('Storage', theme_name)) %>% 
     dplyr::filter(!grepl('Bulk Bricks', theme_name)) %>% 
     dplyr::filter(!grepl('Duplo', theme_name)) %>% 
     dplyr::filter(!grepl('Magnets', theme_name)) %>% 
     dplyr::filter(!grepl('Primo', theme_name)) %>% 
     dplyr::filter(!grepl('Tabletop Games and Puzzles', theme_name)) %>% 
     dplyr::filter(!grepl('LEGO Originals', theme_name)) %>% 
     dplyr::filter(!grepl('Power Functions', theme_name)) %>% 
     dplyr::filter(!grepl('Powered Up', theme_name)) %>% 
     dplyr::filter(!grepl('Books', theme_name)) %>% 
     dplyr::filter(!grepl('Books', theme_name)) %>% 
     dplyr::filter(!grepl('Value', theme_name)) #%>% 
     #dplyr::filter(n > 9) %>% dplyr::arrange(desc(n))

#------------------------------------------------------------------------------
# Groups of models
# ALL SETS.
subset_parts_themes_n10_H<- subset_parts_themes_n10 %>%
    group_by(set_name) %>%
    dplyr::summarize(H= vegan::diversity(table(quantity), 
                 index = "shannon", MARGIN = 1, base = exp(1)))
subset_parts_themes_n10_Simp<- subset_parts_themes_n10 %>%
    group_by(set_name) %>%
    dplyr::summarize(simp= vegan::diversity(table(quantity), "simpson")) 
subset_parts_themes_n10_invSimp<- subset_parts_themes_n10 %>%
    group_by(set_name) %>%
    dplyr::summarize(invSimp= vegan::diversity(table(quantity), 
                 index = "inv")) 
subset_parts_themes_n10_unbSimp<- subset_parts_themes_n10 %>%
    group_by(set_name) %>%
    dplyr::summarize(unbias.simp= vegan::rarefy(table(quantity), 
                     2) - 1) 
# stw_alpha<- Starwars %>%
#     group_by(set_name) %>%
#     dplyr::summarize(alpha= vegan::fisher.alpha(table(quantity))) 
subset_parts_themes_n10_S<- subset_parts_themes_n10 %>%
    group_by(set_name) %>%
    dplyr::summarize(S= vegan::specnumber(table(quantity)))
subset_parts_themes_n10_J<- as.data.frame(cbind(set_name=subset_parts_themes_n10_H$set_name, J= subset_parts_themes_n10_H$H/log(subset_parts_themes_n10_S$S)))
#(J= H/log(S))

#------------------------------------------------------------------------------
# ALL SETS INCLUDED. 
subset_parts_themes_n10_div<- as.data.frame(cbind(
                            all_H, 
                            simp= all_Simp$simp,
                            invSimp= all_invSimp$invSimp,
                            unbias.simp= all_unbSimp$unbias.simp,
                            S= all_S$S,
                            J= all_J$J
                            ))

tt<- as.data.frame(table(subset_parts_themes_n10$set_name))
tt <- rename(tt, "set_name"=Var1, "num_parts"=Freq)
tt<- inner_join(tt, subset_parts_themes_n10[,c(4,5,6,1,2,8,11,12)], by = "set_name",multiple="first")

# allsets_div<-  as.data.frame(cbind(tt,allsets_div))
subset_parts_themes_n10_div <- left_join(tt, subset_parts_themes_n10_div,
  by="set_name") 

subset_parts_themes_n10_div<- subset_parts_themes_n10_div %>% # Filtering out H'=0 records.
        dplyr::filter(H > 0.5)  %>% 
        dplyr::filter(num_parts> 15)

# Plots of H' vs no. parts for sets.
# 
plot(subset_parts_themes_n10_div$num_parts,subset_parts_themes_n10_div$H)
gg <- ggplot(subset_parts_themes_n10_div, aes(x = num_parts, y = H)) +
    geom_point() +
    scale_x_log10() + scale_y_log10() +
    theme_bw()
gg

```


### Network plots

```{r networks, echo=TRUE, fig.height=12, fig.width=12}
# Input edgelist
model75344 <- readr::read_csv(here::here("data/models_adj/75344-1-Boba Fett's Starship Microfighter-el.csv"), show_col_types = FALSE)

model75344.g<- igraph::graph_from_edgelist(as.matrix(cbind(model75344[,c(1,2)])),
                                   directed= F)
#E(model75344.g)$weight<- as.vector(model75344[,4])
wt<- as.vector(model75344[,4])
com <- cluster_leading_eigen(model75344.g)

eb <- cluster_edge_betweenness(model75344.g)
V(model75344.g)$color <- com$membership+1

pal2 <- rainbow(9, alpha=.5)
vertex_attr(model75344.g)

# Plot of parts interactions. Numbers are LEGo part numbers.
# Colors are clusters.
# 
igraph::plot.igraph(model75344.g, layout = layout.fruchterman.reingold, 
     # === vertex
     # === vertex label
     vertex.label.family="Helvetica",
     vertex.label.font= 2,
     vertex.label.cex= 1.25, 
     vertex.label.color= "black",
     vertex.color= adjustcolor(model75344.g$color, alpha.f = .7),
     vertex.label.dist= 0, 
     # === Edge
     edge.color="darkgrey",
     edge.width= 3,
     edge.lty="solid",     
     edge.curved= 0.725,
     # === Other
     margin= c(0,0,0,0), rescale= TRUE
     )

# A 3D plot, interactive. Needs X11.
coords <- layout_with_fr(model75344.g, dim=3)
rglplot(model75344.g, 
        layout=coords,
        vertex.label.color= "black",
        vertex.color= model75344.g$color,
        vertex.label.dist= 2.35,
        vertex.label.degree= pi , 
        edge.width=2                      
     )

```

### Plots with `ggraph`

```{r ggraph_plots, echo=TRUE, message=FALSE, fig.height=12, fig.width=12}
library(ggraph)
require(tidygraph)
wt<- as.vector(model75344[,4])
ggp <- ggraph(model75344.g, 'igraph', algorithm = 'kk') + 
  #geom_edge_link(aes(width = wt), alpha = 0.60, edge_color = "grey20") +
  geom_edge_arc(strength = 0.4) +
  #scale_edge_width_continuous(model75344[,4]) +
  coord_fixed() + 
  scale_edge_alpha('model75344[,4]') +
  geom_node_point(color = 'steelblue', size = 6, alpha= 0.7) +
  ggforce::theme_no_axes() +
  ggtitle("75344-1-Boba Fett's Starship")

# readPNG function to load an image that we have stored on our computer:
my_image <- readPNG("images/75344-1-Boba Fett's Starship.png", native = TRUE)

# Use the inset_element function to add our picture to our ggplot2 graphic:

ggp_image <- ggp +                  # Combine plot & image
  inset_element(p = my_image,
                left = 0.65,
                bottom = 0.1,
                right = 0.9,
                top = 0.45, align_to = 'full', clip=T)
ggp_image                           # Draw combined plot

```
----------------------------------
```{r session-info}
sessionInfo()
```
----------------------------------