---
title: "test_lego_model_network"
author: "Pedro Jordano"
date: "2026-03-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(here)
library(stringr)
library(tidyr)
library(igraph)
library(ggnetwork); library(intergraph)
library(igraph)
library(ggraph)
library(ggplot2)
library(GGally)

library(jsonlite)  # For API
library(purrr)
library(httr)
library(ggimage)
library(grid)
library("png")                      # Load png package
# install.packages("patchwork")       # Install patchwork package
library("patchwork")                # Load patchwork

```

## Introduction

Here is a minimal R workflow that parses this MPD and builds a simple **part–part**
contact network by treating parts that appear in the same submodel as connected. You can
refine the contact criterion later (e.g. geometric distance) if you wish.

## **1. Read and parse the MPD**

Parse an `.ldr` or and `.mpd` file to get the part list.

```{r parse, echo=TRUE}
# 1) Read file
mpd_lines <- readLines(here::here("data/models/4489-1-AT-AT-Mini.mpd"))

# Helper: identify FILE sections
file_idx  <- grep("^0 FILE", mpd_lines)
file_idx_end <- c(file_idx[-1] - 1, length(mpd_lines))

files_df <- tibble(
    file_start = file_idx,
    file_end   = file_idx_end,
    file_header = mpd_lines[file_idx]
) %>%
    mutate(
        file_name = str_trim(str_remove(file_header, "^0 FILE")),
        file_name = str_replace(file_name, "^\\d+\\s+-\\s+", "")  # strip leading "106 - "
    )

files_df

```

## **2. Extract part instances (LDraw type 1 lines)**

You can also embed plots, for example:

```{r parse_1, echo=TRUE}
# Function to parse "1 ..." lines inside a FILE block
parse_file_parts <- function(lines, file_name) {
  # LDraw "1" line: 1 col x y z a b c d e f g h i part.dat or subfile.ldr
  part_lines_idx <- grep("^1\\s", lines)
  if (length(part_lines_idx) == 0) return(NULL)

  # Split by whitespace
  tokens <- str_split(lines[part_lines_idx], "\\s+", simplify = FALSE)

  # Map to data frame
  parts <- lapply(tokens, function(tok) {
    # minimum length check
    if (length(tok) < 15) return(NULL)
    tibble(
      file      = file_name,
      color     = tok[2],
      x         = as.numeric(tok[3]),
      y         = as.numeric(tok[4]),
      z         = as.numeric(tok[5]),
      a         = as.numeric(tok[6]),
      b         = as.numeric(tok[7]),
      c         = as.numeric(tok[8]),
      d         = as.numeric(tok[9]),
      e         = as.numeric(tok[10]),
      f         = as.numeric(tok[11]),
      g         = as.numeric(tok[12]),
      h         = as.numeric(tok[13]),
      i         = as.numeric(tok[14]),
      part_raw  = paste(tok[15:length(tok)], collapse = " ")
    )
  })

  bind_rows(parts)
}

# Apply to each FILE block
parts_all <- purrr::map2_dfr(
  files_df$file_start,
  files_df$file_end,
  ~{
    block_lines <- mpd_lines[.x:.y]
    fname <- files_df$file_name[files_df$file_start == .x]
    parse_file_parts(block_lines, fname)
  }
)

parts_all

```

This gives a table with one row per placed element (including submodels like
`106 - car.ldr`, `106 - Minifig.ldr`, and `106 - wheel assembly.ldr`).

## **3. Decide what counts as a “part” and “connection”**

For a first simple network, you can:

-   Represent each **part file** (`*.dat`, `*.ldr`) as a node.

-   Add an edge between two parts if they appear together in the same FILE (submodel).

Below I create a co‑occurrence network at the level of **part_raw** (so `3004.dat` etc.
are nodes). You can easily switch to unique instance IDs if you want a multigraph.

```{r parts_as_nodes, echo=TRUE}
# 3) Clean part name (strip leading "106 - " from submodels, keep .dat etc.)
parts_all <- parts_all %>%
  mutate(
    part = str_trim(part_raw)
  )

# 4) Build co-occurrence edges:
# For each FILE, connect all distinct part types that co-occur
edges_df <- parts_all %>%
  group_by(file) %>%
  summarise(parts = list(unique(part)), .groups = "drop") %>%
  mutate(
    edges = purrr::map(parts, ~{
      p <- .x
      if (length(p) < 2) return(NULL)
      # all unordered pairs
      comb <- t(combn(p, 2))
      tibble(from = comb[,1], to = comb[,2])
    })
  ) %>%
  select(edges) %>%
  unnest(edges)

edges_df

```

## **4. Build igraph object and analyze**

```{r build_graph, echo=TRUE}
# 5) Create graph
g <- graph_from_data_frame(edges_df, directed = FALSE)

# Basic stats
print(g)
degree(g)

# # Example: plot
# plot(
#   g,
#   vertex.size = 5,
#   vertex.label.cex = 0.6,
#   vertex.label.dist = 0.5,
#   main = "LEGO parts co-occurrence network"
# )

# Using ggraph (Recommended)
# ggraph extends ggplot2 for networks and works directly with igraph.

ggp <- ggraph(g, layout = "fr") +  # Fruchterman-Reingold layout
# geom_edge_link(aes(width = 0.0005), alpha = 0.2) +
#  geom_edge_link(aes(alpha= 0.3)) +
#  geom_edge_link(width = 0.1, alpha = 0.7) +
  geom_edge_bend0(width = 0.18, alpha = 0.7) +
  geom_node_point(aes(size = degree(g)), color = "blue", alpha= 0.65) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_graph() +
  labs(title = "Network with 4489-1-AT-AT-Mini")

# readPNG function to load an image that we have stored on our computer:
my_image <- readPNG("images/4489-1-AT-AT-Mini.png", native = TRUE)

# Use the inset_element function to add our picture to our ggplot2 graphic:

ggp_image <- ggp +                  
  inset_element(p = my_image,
                left = 0.4,
                bottom = 1,
                right = 0.3,
                top = 0.8, 
                align_to = 'full', 
                clip = TRUE)
ggp_image                           # Draw combined plot

```

## **5. Optional: distinguish submodels vs primitive parts**

We can tag nodes as “submodel” when they are `.ldr` vs `.dat`, which might help link to
assembly hierarchy.

```{r submodels, echo=TRUE}
node_attr <- tibble(
  part = V(g)$name
) %>%
  mutate(
    is_submodel = str_detect(part, "\\.ldr$"),
    is_primitive = str_detect(part, "\\.dat$")
  )

V(g)$is_submodel  <- node_attr$is_submodel[match(V(g)$name, node_attr$part)]
V(g)$is_primitive <- node_attr$is_primitive[match(V(g)$name, node_attr$part)]

```

We may prefer the network at the level of **individual brick instances** (each physical
brick a node) with edges defined by spatial proximity, or is this part‑type co‑occurrence
representation what we want?

### See file `LEGO_model_networks_ind-parts.Rmd`

# Network visualization with part images

Below is an end‑to‑end pattern tailored to a Rebrickable parts table and LEGO‑network workflow.


## 1. Download and cache images locally
Assume you already have a parts data.frame like parts_4489 from get_set_parts():
parts_4489 columns: part_num, part_name, color_id, color_name, img_url, quantity

```{r dowload_part_images, echo=TRUE}
# CSV download with function get_set_parts.R: 
# https://cdn.rebrickable.com/media/downloads/sets.csv.zip?
api_key <- "40c6692458e8b4c27074486ab114d6cc"
source(here::here("./code/connectivity_workflow/functions/get_set_parts.R"))

# Dataframe with parts list for a set.
parts_4489 <- get_set_parts("4489-1", api_key)
head(parts_4489)
parts_4489 <- parts_4489 %>%
    dplyr::mutate(
        part_num_trim = sub("\\.dat$", "", part_num)
    )
parts_4489new <- parts_4489 %>%
    dplyr::filter(part_num_trim %in% V(g)$part_id)
#

download_part_images <- function(urls, destdir = "images/part_images") {
    dir.create(destdir, showWarnings = FALSE)
    paths <- file.path(destdir, basename(urls))
    
    mapply(
        function(u, p) {
            if (!file.exists(p) && !is.na(u) && nzchar(u)) {
                try(download.file(u, p, mode = "wb"), silent = TRUE)
            }
        },
        urls, paths
    )
    
    paths
}

parts_4489$img_path <- download_part_images(parts_4489$img_url, 
                                            destdir = "images/part_images")

# Subsetting the parts file to just unique part types (no color or duplications).
parts_4489_unique <- parts_4489 %>%
    distinct(part_num, .keep_all = TRUE)

# NO parts_4489_unique$part_num en V(g)$part_id: 44567,30000,4592,4593,4489 - Head.ldr,
exclude_parts <- c("44567", "30000", "4592", "4593", "4489 - Head.ldr")

parts_4489_filtered <- parts_4489_unique %>%
    dplyr::filter(!part_num %in% exclude_parts)

```

Basename(urls) keeps names simple; we can choose more informative filenames if we want 
(e.g., paste `part_num` and `color_id`).

After this step, `parts_4489_filtered$img_path` points to local PNG files.


## 2. Attach image paths to igraph nodes

Assume:
    The igraph object is g.
Each node name equals a part_num (e.g., "3001", "3024").
Ensure node names are part_nums
V(g)$name already something like "3001", "3024", ...
Assume V(g)$name are part_ids like "3001"; map 20 nodes

```{r iamges_to_nodes,echo=TRUE}
# This sets url of part images to the nodes identified by their part number.

V(g)$part_id <- sub("\\.dat$", "", V(g)$name) #V(g)$name  # Or your mapping
V(g)$color_id <- 7         # parts_4489new$color_id  
                           # Light_grey example; get from parts_json$results$color_id
# V(g)$image <-parts_4489_filtered$img_url

# 1) Build node table from graph vertices
nodes <- tibble::tibble(part_id = V(g)$part_id)

# 2) Left-join image URLs from filtered parts by part number
nodes <- nodes %>%
  left_join(
    parts_4489_unique %>%
    dplyr::select(part_num, img_path),
    by = c("part_id" = "part_num")
  )

# 3) Check alignment
nrow(nodes)          # should be 31
vcount(g)            # should also be 31

# 4) Assign to graph
V(g)$img_url <- nodes$img_path

```


## 3. Compute layout and node data frame and plot with Rebrickable URLs

```{r plotting, echo=TRUE}
# set.seed(1)
# coords <- layout_with_fr(g) * 1.5
# 
# nodes_plot <- data.frame(
#     x       = coords[,1],
#     y       = coords[,2],
#     name    = V(g)$part_id,  # or V(g)$name
#     img_url = V(g)$img_url
# )
# 
# # 3) Test 
# 
# ggraph(g, layout = "manual", x = coords[,1], y = coords[,2]) +
#     geom_edge_link(width = 0.1, alpha = 0.8) +
#     geom_node_point(size = 0, alpha = 0) +
#     geom_image(
#         data = nodes_plot,
#         aes(x = x, y = y, image = img_url), # This grabs the image from Rebrickable url.
#         size = 0.06,
#         inherit.aes = FALSE
#     ) +
#     geom_node_text(aes(label = name),
#                    vjust = 2.25,
#                    size = 4,
#                    color = "black") +
#     theme_graph() +
#     labs(title = "LEGO Parts Network (4489-1 AT-AT-Mini)")

```

Now let's each node having a `V(g)$img_path` pointing to a local file.

## 4. Plot with ggraph + ggimage using local files

### Key points:
In `geom_image()`, we can pass the vector directly or construct a tibble with node attributes and map `aes(image = img_path)`.

`size` is relative to the plot; adjust (e.g. 0.07–0.12) to fit ~20 nodes.

Also, I use a safer filename scheme. To avoid clashes when two URLs share the same basename:

```{r safer_scheme, echo=TRUE}
download_part_images <- function(part_num, color_id, urls, 
                                 destdir = "images/part_images/") {
    dir.create(destdir, showWarnings = FALSE)
    filenames <- paste0(part_num,".png")
    paths <- file.path(destdir, filenames)
    mapply(
        function(u, p) {
            if (!file.exists(p) && !is.na(u) && nzchar(u)) {
                try(download.file(u, p, mode = "wb"), silent = TRUE)
            }
        },
        urls, paths
    )
    
    paths
}

parts_4489$img_path <- download_part_images(
    part_num = parts_4489$part_num,
    color_id = parts_4489$color_id,
    urls     = parts_4489$img_url,
    destdir  = "images/part_images/"
)

# parts_4489$img_path <- download_part_images(parts_4489$img_url, destdir = "images")

#Check part names include .dat
#parts_4489 <- parts_4489 %>%
#    dplyr::mutate(part_num2= paste0(part_num,".dat"))

# download.file(V(g)$image, destdir = "part_imgs/")
V(g)$loc.image <- paste0("images/part_images/", V(g)$part_id, ".png")


# 3. Plot with ggraph + ggimage using local files
# 1) Layout
coords <- layout_with_fr(g) * 1.5

# 2) Node data frame with coords + attributes
nodes <- data.frame(
    x        = coords[, 1],
    y        = coords[, 2],
    name     = V(g)$part_id,  # or V(g)$name
    img_path = V(g)$loc.image,
    degree   = degree(g)
)

# 3) Plot: ggraph uses g + manual layout; geom_image uses nodes with its own aes
ggp<- ggraph(g, layout = "manual", x = coords[,1], y = coords[,2]) +
  geom_edge_bend0(width = 0.18, alpha = 0.7) +
# geom_node_point(size = 0, alpha = 0) +
  geom_node_point(aes(size = degree(g)), color = "blue", alpha= 0.65) +
  geom_edge_link(width = 0.1, alpha = 0.8) +
  geom_image(
        data = nodes,
        aes(x = x, y = y, image = img_path), use_cache=FALSE,
        size = 0.05,                 # fixed, reasonable size
        inherit.aes = FALSE
    ) +
#  geom_node_text(aes(label = name), repel = TRUE) +
  geom_node_text(aes(label = name),
                   vjust = 4.25,
                   size = 4,
                   color = "black") +
    theme_graph() +
    labs(title = "LEGO Parts Network (4489-1 AT-AT-Mini)")

# readPNG function to load an image that we have stored on our computer:
img_array <- readPNG("images/4489-1-AT-AT-Mini.png", native = TRUE)
my_image_grob <- rasterGrob(img_array, interpolate = TRUE)

# If still see overlap, shrink further:
# decrease width: increase left to 0.1 or decrease right to 0.3
# decrease height: increase bottom to 0.7
ggp_image <- ggp +
  inset_element(
    my_image_grob,
    left   = 0,
    bottom = 0.8,
    right  = 1,
    top    = 0.98,
    align_to = "full",
    clip     = TRUE
  )
ggp_image <- ggp +
  inset_element(
    my_image_grob,
    left   = 0.15,  # 15% from left
    right  = 0.15,  # 65% from left -> 30% width
    bottom = 0.90,  # 35% from bottom
    top    = 0.6,  # 65% from bottom -> 30% height
    align_to = "full",
    clip     = F
  )

ggp_image                           # Draw combined plot

# Key points:
#    nodes must contain columns literally named x and y.

# In geom_image(...) we set data = nodes and aes(x = x, y = y, image = img_path, size = degree) so all needed aesthetics exist in that data.

# inherit.aes = FALSE stops it from looking for x/y in the ggraph layout environment, so there is no ambiguity.

# size is relative to the plot; adjust (e.g. 0.07–0.12) to fit ~20 nodes.
# Using ggraph (Recommended)
# ggraph extends ggplot2 for networks and works directly with igraph.

```

-----------------------------------------------------------------------------------------------

# With function
# CHECK
```{r plot_lego_set_connectivity, echo=TRUE}
#Sys.setenv(REBRICKABLE_KEY = "40c6692458e8b4c27074486ab114d6cc")

#g_689 <- plot_lego_set_connectivity("689-1")

```

-----------------------------------------------------------------------------------------------
## Session Info

```{r session, echo=TRUE}
session_info()

```

