---
title: "test_lego_model_network"
author: "Pedro Jordano"
date: "2026-03-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(here)
library(stringr)
library(tidyr)
library(igraph)
library(ggnetwork); library(intergraph)
library(igraph)
library(ggraph)
library(ggplot2)
library(GGally)

library(jsonlite)  # For API
library(purrr)
library(httr)
library(ggimage)
library(grid)
library("png")                      # Load png package
# install.packages("patchwork")       # Install patchwork package
library("patchwork")                # Load patchwork

```

## Introduction

Here is a minimal R workflow that parses this MPD and builds a simple **part–part**
contact network by treating parts that appear in the same submodel as connected. We can
refine the contact criterion later (e.g. geometric distance) if needed.

## **1. Read and parse the MPD**

Parse an `.ldr` or and `.mpd` file to get the part list.

```{r parse, echo=TRUE}
# 1) Read a .mdp file for the selected LEGO set.
set_id<- "8008-1" ###### <<<<<------- Input set code here.
mpd_lines <- readLines(here::here(paste0("data/models/starwars/",set_id,".mpd")))

# Helper: extract setname.
source(here::here("functions/extract_mpd_set_name.R"))
setname<- extract_mpd_set_name(here::here("data/models/starwars/7110-1.mpd"))

# readPNG function to load an image of the set that we have stored on our computer:
# For later use as inset in a network graph.
my_image <- readPNG(here::here(paste0("images/",setname,"png"), 
                               native = TRUE))

# Helper: identify FILE sections
file_idx  <- grep("^0 FILE", mpd_lines)
file_idx_end <- c(file_idx[-1] - 1, length(mpd_lines))

files_df <- tibble(
    file_start = file_idx,
    file_end   = file_idx_end,
    file_header = mpd_lines[file_idx]
) %>%
    mutate(
        file_name = str_trim(str_remove(file_header, "^0 FILE")),
        file_name = str_replace(file_name, "^\\d+\\s+-\\s+", "")  # strip leading "106 - "
    )

files_df

```

## **2. Extract part instances (LDraw type 1 lines)**

```{r parse_1, echo=TRUE}
# Function to parse "1 ..." lines inside a FILE block
parse_file_parts <- function(lines, file_name) {
  # LDraw "1" line: 1 col x y z a b c d e f g h i part.dat or subfile.ldr
  part_lines_idx <- grep("^1\\s", lines)
  if (length(part_lines_idx) == 0) return(NULL)

  # Split by whitespace
  tokens <- str_split(lines[part_lines_idx], "\\s+", simplify = FALSE)

  # Map to data frame
  parts <- lapply(tokens, function(tok) {
    # minimum length check
    if (length(tok) < 15) return(NULL)
    tibble(
      file      = file_name,
      color     = tok[2],
      x         = as.numeric(tok[3]),
      y         = as.numeric(tok[4]),
      z         = as.numeric(tok[5]),
      a         = as.numeric(tok[6]),
      b         = as.numeric(tok[7]),
      c         = as.numeric(tok[8]),
      d         = as.numeric(tok[9]),
      e         = as.numeric(tok[10]),
      f         = as.numeric(tok[11]),
      g         = as.numeric(tok[12]),
      h         = as.numeric(tok[13]),
      i         = as.numeric(tok[14]),
      part_raw  = paste(tok[15:length(tok)], collapse = " ")
    )
  })

  bind_rows(parts)
}

# Apply to each FILE block
parts_all <- purrr::map2_dfr(
  files_df$file_start,
  files_df$file_end,
  ~{
    block_lines <- mpd_lines[.x:.y]
    fname <- files_df$file_name[files_df$file_start == .x]
    parse_file_parts(block_lines, fname)
  }
)

parts_all

```

This gives a table with one row per placed element (including submodels like
`106 - car.ldr`, `106 - Minifig.ldr`, and `106 - wheel assembly.ldr`).

## **3. Decide what counts as a “part” and “connection”**

For a first simple network, you can:

-   Represent each **part file** (`*.dat`, `*.ldr`) as a node.

-   Add an edge between two parts if they appear together in the same FILE (submodel).

Below I create a co‑occurrence network at the level of **part_raw** (so `3004.dat` etc.
are nodes). You can easily switch to unique instance IDs if you want a multigraph.

```{r parts_as_nodes, echo=TRUE}
# 3) Clean part name (strip leading "106 - " from submodels, keep .dat etc.)
parts_all <- parts_all %>%
  mutate(
    part = str_trim(part_raw)
  )

# 4) Build co-occurrence edges:
# For each FILE, connect all distinct part types that co-occur
edges_df <- parts_all %>%
  group_by(file) %>%
  summarise(parts = list(unique(part)), .groups = "drop") %>%
  mutate(
    edges = purrr::map(parts, ~{
      p <- .x
      if (length(p) < 2) return(NULL)
      # all unordered pairs
      comb <- t(combn(p, 2))
      tibble(from = comb[,1], to = comb[,2])
    })
  ) %>%
  select(edges) %>%
  unnest(edges)

edges_df

```

## **4. Build igraph object and analyze**

```{r build_graph, echo=TRUE}
# 5) Create graph
g <- graph_from_data_frame(edges_df, directed = FALSE)

# Basic stats
print(g)
degree(g)

# # Example: plot
# plot(
#   g,
#   vertex.size = 5,
#   vertex.label.cex = 0.6,
#   vertex.label.dist = 0.5,
#   main = "LEGO parts co-occurrence network"
# )

# Using ggraph (Recommended)
# ggraph extends ggplot2 for networks and works directly with igraph.

ggp <- ggraph(g, layout = "fr",
              niter= 300) +  # Fruchterman-Reingold layout
# geom_edge_link(aes(width = 0.0005), alpha = 0.2) +
#  geom_edge_link(aes(alpha= 0.3)) +
#  geom_edge_link(width = 0.1, alpha = 0.7) +
  geom_edge_bend0(width = 0.18, alpha = 0.7) +
  geom_node_point(aes(size = degree(g)), color = "blue", alpha= 0.65) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_graph() +
  labs(title = print(setname))

# Pre-compute bounds for compact area (adjust based on your g)
# n_nodes <- vcount(g)
# area_size <- sqrt(n_nodes) * 3  # Scales with graph size
# minC <- rep(-area_size/2, 2)
# maxC <- rep(area_size/2, 2)

# More ompact arrangement.
# ggp <- ggraph(g, 
#               layout = "fr", 
#               niter = 300) +  # Fewer iterations = tighter/faster
#   geom_edge_bend0(width = 0.18, alpha = 0.7) +
#   geom_node_point(aessize = degree(g), #(size = pmin(degree(g), 8)), 
#                   color = "blue", alpha = 0.7) +
#   geom_node_text(aes(label = name), repel = TRUE) +
# #  geom_node_text(aes(label = ifelse(degree(g) > 2, name, "")), 
# #                 repel = TRUE, size = 2.5) +
#   theme_graph() +
#   labs(title = setname) +
#   coord_fixed(0.6)

# Use the inset_element function to add our picture to our ggplot2 graphic:

ggp_image <- ggp +                  
  inset_element(p = my_image,
                left = 0.05,
                bottom = 0.5,
                right = 0.19,
                top = 0.8, 
                align_to = 'full', 
                clip = TRUE)
ggp_image                           # Draw combined plot

```

## **5. Optional: distinguish submodels vs primitive parts**

We can tag nodes as “submodel” when they are `.ldr` vs `.dat`, which might help link to
assembly hierarchy.

```{r submodels, echo=TRUE}
node_attr <- tibble(
  part = V(g)$name
) %>%
  mutate(
    is_submodel = str_detect(part, "\\.ldr$"),
    is_primitive = str_detect(part, "\\.dat$")
  )

V(g)$is_submodel  <- node_attr$is_submodel[match(V(g)$name, node_attr$part)]
V(g)$is_primitive <- node_attr$is_primitive[match(V(g)$name, node_attr$part)]

```

We may prefer the network at the level of **individual brick instances** (each physical
brick a node) with edges defined by spatial proximity, or is this part‑type co‑occurrence
representation what we want?

### See file `LEGO_model_networks_ind-parts.Rmd`

# Network visualization with part images

Below is an end‑to‑end pattern tailored to a Rebrickable parts table and LEGO‑network workflow.

## 1. Download and cache images locally
Assume we already have a parts `data.frame` like `parts_4489` from function `get_set_parts()`:
`parts_4489` columns: `part_num`, `part_name`, `color_id`, `color_name`, `img_url`, `quantity`

```{r dowload_part_images, echo=TRUE}
## 1. Download and cache images locally
# Assume we already have a parts `data.frame` like `parts_4489` from function 
# `get_set_parts()`:
# `parts_4489` columns: `part_num`, `part_name`, `color_id`, `color_name`, 
# `img_url`, `quantity`
#######################################
target_set<- c("4489-1")
#######################################
#
# CSV download with function get_set_parts.R: 
# https://cdn.rebrickable.com/media/downloads/sets.csv.zip?
api_key <- "40c6692458e8b4c27074486ab114d6cc"
source(here::here("./code/connectivity_workflow/functions/get_set_parts.R"))

# Dataframe with parts list for a set.
target_set_parts <- get_set_parts(target_set, api_key)
head(target_set_parts) # part_num contains the parts numbers.

# Add local url for part images of dataset. 
source(here::here("functions/get_LEGO_part_images_for_set.R"))
target_set_parts <- get_LEGO_part_images_for_set(target_set)

target_set_parts <- target_set_parts %>%
    distinct(part_num, .keep_all = TRUE)

```

Basename(urls) keeps names simple; we can choose more informative filenames if we want 
(e.g., paste `part_num` and `color_id`).

After this step, `parts_4489_filtered$img_path` points to local PNG files.


## 2. Compute layout and node data frame and plot with Rebrickable URLs, using ggraph + ggimage using local files

Assume:
    The igraph object is g.
Each node name equals a part_num (e.g., "3001", "3024").
Ensure node names are part_nums
V(g)$name already something - like "3001", "3024", ...
Assume V(g)$name are part_ids like "3001"; map 20 nodes

Now let's each node having a `V(g)$img_path` pointing to a local file.

### Key points:
In `geom_image()`, we can pass the vector directly or construct a tibble with node attributes and map `aes(image = img_path)`.

`size` is relative to the plot; adjust (e.g. 0.07–0.12) to fit ~20 nodes.

Also, I use a safer filename scheme. To avoid clashes when two URLs share the same basename:

```{r iamges_to_nodes,echo=TRUE}
# This sets url of part images to the nodes identified by their part number.

V(g)$part_id <- sub("\\.dat$", "", V(g)$name) #V(g)$name  # Or your mapping
V(g)$color_id <- 7         # parts_4489new$color_id  
                           # Light_grey example; get from parts_json$results$color_id

# 1) Build node table from graph vertices
nodes <- tibble::tibble(part_id = V(g)$part_id)

# 2) Left-join image URLs from filtered parts by part number
nodes <- nodes %>%
  left_join(
    target_set_parts %>%
    dplyr::select(part_num, img_local_path),
    by = c("part_id" = "part_num")
  )

# 3) Check alignment
nrow(nodes)          # should be 31
vcount(g)            # should also be 31

# download.file(V(g)$image, destdir = "part_imgs/")
V(g)$loc.image <- paste0("images/part_images/", V(g)$part_id, ".png")

##############
# Plottng.
##############
source(here::here("functions/setup_graph_plot.R"))
setup_graph_plot(target_set)

coords <- layout_with_fr(g) * 1.5
# diag_coords_OK<- coords

# Actual plot.
source(here::here("functions/plot_bend_net.R"))
plot_bend_net(g, coords)

```

-----------------------------------------------------------------------------------------------
## Session Info

```{r session, echo=TRUE}
session_info()

```

