---
title: "test_lego_model_network"
author: "Pedro Jordano"
date: "2026-03-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(here)
library(stringr)
library(tidyr)
library(igraph)
library(ggnetwork); library(intergraph)
library(igraph)
library(ggraph)
library(ggplot2)
library(GGally)

library(jsonlite)  # For API
library(purrr)
library(httr)
library(ggimage)

```



## 6. Network at the level of individual brick instances (each physical brick a node)

To get a network where each node is a physical brick instance, we treat every LDraw
`1` line as a unique element and then define edges by geometric proximity in world
coordinates.

Below is a self‑contained R script that:

-   Parses the MPD, including nested submodels (`.ldr` inside `.ldr`).

-   Computes global coordinates for every brick instance.

-   Builds a simple contact network: edge if distance between brick centroids is below a
    chosen threshold (e.g. 20 LDraw units ≈ 2 studs).


```{r indiv_bricks_network, echo=TRUE}
library(dplyr)
library(stringr)
library(purrr)
library(tidyr)
library(igraph)

#-----------------------------
# 1. Read MPD and split FILEs
#-----------------------------
mpd_lines <- readLines(here::here("connectivity_workflw/106-Unicef_Truck.mpd"))

file_idx      <- grep("^0 FILE", mpd_lines)
file_idx_end  <- c(file_idx[-1] - 1, length(mpd_lines))

files_df <- tibble(
  file_start = file_idx,
  file_end   = file_idx_end,
  file_header = mpd_lines[file_idx]
) %>%
  mutate(
    file_name_raw = str_trim(str_remove(file_header, "^0 FILE")),
    # e.g. "106 - car.ldr" -> "car.ldr"
    file_name = file_name_raw %>%
      str_replace("^\\d+\\s+-\\s+", "") %>%
      str_trim()
  )

#-----------------------------------
# 2. Parse part instances per FILE
#-----------------------------------
parse_file_parts <- function(lines, file_name) {
  idx <- grep("^1\\s", lines)
  if (length(idx) == 0) return(NULL)

  tokens_list <- str_split(lines[idx], "\\s+", simplify = FALSE)

  rows <- map(tokens_list, function(tok) {
    if (length(tok) < 15) return(NULL)
    tibble(
      file      = file_name,
      color     = tok[2],
      x         = as.numeric(tok[3]),
      y         = as.numeric(tok[4]),
      z         = as.numeric(tok[5]),
      a         = as.numeric(tok[6]),
      b         = as.numeric(tok[7]),
      c         = as.numeric(tok[8]),
      d         = as.numeric(tok[9]),
      e         = as.numeric(tok[10]),
      f         = as.numeric(tok[11]),
      g         = as.numeric(tok[12]),
      h         = as.numeric(tok[13]),
      i         = as.numeric(tok[14]),
      part_raw  = paste(tok[15:length(tok)], collapse = " ")
    )
  })

  bind_rows(rows)
}

parts_local <- map2_dfr(
  files_df$file_start,
  files_df$file_end,
  ~{
    block_lines <- mpd_lines[.x:.y]
    fname <- files_df$file_name[files_df$file_start == .x]
    parse_file_parts(block_lines, fname)
  }
)

# Give each local instance an ID within its FILE
parts_local <- parts_local %>%
  group_by(file) %>%
  mutate(local_id = row_number()) %>%
  ungroup()

#-----------------------------------
# 3. Build file lookup (submodels)
#-----------------------------------
file_lookup <- files_df %>%
  select(file_name) %>%
  mutate(is_submodel = TRUE)

# A part is a submodel if its name (after cleaning) matches a FILE name
parts_local <- parts_local %>%
  mutate(
    part = str_trim(part_raw),
    # also strip "106 - " from part if present
    part_clean = part %>%
      str_replace("^\\d+\\s+-\\s+", "") %>%
      str_trim()
  ) %>%
  left_join(file_lookup, by = c("part_clean" = "file_name")) %>%
  mutate(is_submodel = ifelse(is.na(is_submodel), FALSE, is_submodel))

#-----------------------------------
# 4. Recursive expansion to world coords
#-----------------------------------
# Represent transforms as 3x3 matrix + translation vector
mat_from_row <- function(row) {
  matrix(c(row$a, row$b, row$c,
           row$d, row$e, row$f,
           row$g, row$h, row$i),
         nrow = 3, byrow = TRUE)
}

# Depth-first traversal: expand submodels to primitive brick instances
expand_file <- function(file_name, parent_mat, parent_trans, parent_path) {
  # parent_mat: 3x3 matrix
  # parent_trans: length-3 numeric
  # parent_path: character string to track hierarchy

  df <- parts_local %>% filter(file == file_name)

  if (nrow(df) == 0) return(NULL)

  res_list <- vector("list", nrow(df))

  for (k in seq_len(nrow(df))) {
    row <- df[k, ]

    # local transform of this instance
    mat_local <- mat_from_row(row)
    trans_local <- c(row$x, row$y, row$z)

    # compose transform: global = parent_mat %*% local + parent_trans
    mat_global <- parent_mat %*% mat_local
    trans_global <- as.numeric(parent_mat %*% trans_local + parent_trans)

    this_path <- paste0(parent_path, "/", row$file, "[", row$local_id, "]")

    if (isTRUE(row$is_submodel)) {
      # recurse into submodel
      sub_expanded <- expand_file(
        file_name   = row$part_clean,
        parent_mat  = mat_global,
        parent_trans = trans_global,
        parent_path = this_path
      )
      res_list[[k]] <- sub_expanded
    } else {
      # primitive brick instance: store its global transform
      res_list[[k]] <- tibble(
        instance_id = paste0(this_path, ":", row$part),
        file_origin = row$file,
        part        = row$part,
        color       = row$color,
        x_global    = trans_global[1],
        y_global    = trans_global[2],
        z_global    = trans_global[3],
        # flatten rotation matrix for later if needed
        a = mat_global[1,1], b = mat_global[1,2], c = mat_global[1,3],
        d = mat_global[2,1], e = mat_global[2,2], f = mat_global[2,3],
        g = mat_global[3,1], h = mat_global[3,2], i = mat_global[3,3]
      )
    }
  }

  bind_rows(res_list)
}

# Start from main model (first FILE) – here "main.ldr"
root_file <- files_df$file_name[1]

world_parts <- expand_file(
  file_name   == root_file,
  parent_mat  == diag(3),
  parent_trans == c(0, 0, 0),
  parent_path == root_file
)

# world_parts now has one row per physical brick instance in world coordinates. [file:1]
world_parts

#-----------------------------------
# 5. Build contact network between instances
#-----------------------------------
# Simple rule: connect two bricks if their centroids are closer than a threshold.
# LDraw units: 20 LDU ≈ 1 LEGO plate, 24 LDU ≈ 1 stud height; adjust threshold as desired.
dist_threshold <- 25  # tune as needed

coords <- world_parts %>%
  select(instance_id, x_global, y_global, z_global)

n <- nrow(coords)

# pairwise distances (naive O(n^2) – fine for small model)
ij <- t(combn(seq_len(n), 2))
dx <- coords$x_global[ij[,1]] - coords$x_global[ij[,2]]
dy <- coords$y_global[ij[,1]] - coords$y_global[ij[,2]]
dz <- coords$z_global[ij[,1]] - coords$z_global[ij[,2]]
dist <- sqrt(dx^2 + dy^2 + dz^2)

contact_idx <- which(dist <= dist_threshold)

edges_instances <- tibble(
  from = coords$instance_id[ij[contact_idx, 1]],
  to   = coords$instance_id[ij[contact_idx, 2]],
  distance = dist[contact_idx]
)

edges_instances

#-----------------------------------
# 6. Create igraph and do basic analysis
#-----------------------------------
g_inst <- graph_from_data_frame(edges_instances, vertices = world_parts, directed = FALSE)

g_inst
degree(g_inst)

plot(
  g_inst,
  vertex.size = 4,
  vertex.label = NA,
  main = "LEGO brick-instance contact network"
)

```

This gives us a brick‑instance graph where each node is one physical element in the
assembled model, and edges approximate direct contacts based on centroid distance in LDraw
coordinates.

We now want to restrict “contacts” to bricks that are on the same layer (e.g. similar `yy`)
to better approximate actual stud–tube connections, rather than generic spatial proximity.

## 7. Restrict “contacts” to bricks that are on the same layer (e.g. similar `yy`)

We now add a vertical “layer” constraint by only allowing contacts between bricks whose
global yy coordinates are within a small tolerance, and then apply the distance threshold
in the horizontal plane or in full 3D.

Assuming we already have `world_parts` as in the previous script, here is a modified
contact‑network section.

### **1. Define layer tolerance and distance metric**

```{r tolerance, echo=TRUE}
# world_parts: one row per brick instance, with x_global, y_global, z_global

layer_tol      <- 5    # max |Δy| for bricks to be considered on same layer
dist_threshold <- 25   # max 3D distance (or horizontal) for contact

```

### **2. Compute pairwise contacts with layer restriction**

```{r pairw_contacts, echo=TRUE}
coords <- world_parts %>%
  select(instance_id, x_global, y_global, z_global)

n  <- nrow(coords)
ij <- t(combn(seq_len(n), 2))

dx <- coords$x_global[ij[,1]] - coords$x_global[ij[,2]]
dy <- coords$y_global[ij[,1]] - coords$y_global[ij[,2]]
dz <- coords$z_global[ij[,1]] - coords$z_global[ij[,2]]

# 3D distance (we could also use horizontal only: sqrt(dx^2 + dz^2))
dist3d <- sqrt(dx^2 + dy^2 + dz^2)

same_layer <- abs(dy) <= layer_tol
close_enough <- dist3d <= dist_threshold

contact_idx <- which(same_layer & close_enough)

edges_instances <- tibble(
  from      = coords$instance_id[ij[contact_idx, 1]],
  to        = coords$instance_id[ij[contact_idx, 2]],
  distance  = dist3d[contact_idx],
  dy        = dy[contact_idx]
)

edges_instances

```

## **3. Build the instance‑level graph**

```{r instance_level_graph, echo=TRUE}
g_inst <- graph_from_data_frame(
  edges_instances,
  vertices = world_parts,
  directed = FALSE
)

g_inst
degree(g_inst)

plot(
  g_inst,
  vertex.size = 4,
  vertex.label = NA,
  main = "Brick-instance contact network (same-layer restriction)"
)

```

We can tune `layer_tol` to match the expected plate/brick heights in LDraw units; for
example, for only strictly coplanar bricks we set it close to 0, while a slightly
larger window will capture small numerical offsets from rotations.

Now classify layers automatically (e.g. cluster `yy` values into discrete
levels and then restrict contacts to within a level).

```{r classify_layers, echo=TRUE}
node_attr <- tibble(
    part = V(g)$name ) %>% 
    mutate(is_submodel = str_detect(part,"\\.ldr$"), 
           is_primitive = str_detect(part, "\\.dat$"))

V(g)$is_submodel <- node_attr$is_submodel[match(V(g)$name, 
                                                node_attr$part)]
V(g)$is_primitive <- node_attr$is_primitive[match(V(g)$name, 
                                                  node_attr$part)]

```


That snippet tags graph nodes as submodels (`.ldr`) or primitive parts (`.dat`); with
brick‑instance nodes you should adapt it to use instance attributes instead of
`V(g)$name`.

For the **part‑type** network (earlier `g`):

```         
r
```

\`library(stringr) library(tibble) library(dplyr)

node_attr \<- tibble( part = V(g)$name
) %>%
  mutate(
    is_submodel  = str_detect(part, "\\.ldr$"), is_primitive = str_detect(part,
"\\.dat\$") )

V(g)$is_submodel  <- node_attr$is_submodel[match(V(g)$name, node_attr$part)]
V(g)$is_primitive <- node_attr$is_primitive[match(V(g)$name, node_attr$part)]\`

For the **brick‑instance** network (`g_inst`, vertices from `world_parts`), a more useful
version is:

```         
r
```

\`node_attr_inst \<- world_parts %\>% transmute( instance_id, part, color, is_submodel =
str_detect(part, "\\.ldr$"),
    is_primitive = str_detect(part, "\\.dat$") )

V(g_inst)$part        <- node_attr_inst$part[match(V(g_inst)$name, node_attr_inst$instance_id)]
V(g_inst)$color       <- node_attr_inst$color[match(V(g_inst)$name, node_attr_inst$instance_id)]
V(g_inst)$is_submodel <- node_attr_inst$is_submodel[match(V(g_inst)$name, node_attr_inst$instance_id)]
V(g_inst)$is_primitive<- node_attr_inst$is_primitive[match(V(g_inst)$name, node_attr_inst$instance_id)]\`

This way we keep your **instance‑level** network but still know which LEGO element file
each node comes from.

Is your goal to eventually collapse instance‑level contacts into a part‑type interaction
network (e.g. a weighted projection of brick contacts)?

eventually collapse instance‑level contacts into a part‑type interaction network (e.g. a
weighted projection of brick contacts)

You can project the brick‑instance contact network onto part types by grouping edges by
the pair of part files and counting how many instance‑level contacts each pair has.

## **1. Start from instance‑level data**

Assume you already have:

-   `world_parts`: one row per instance, with `instance_id` and `part`.

-   `edges_instances`: contacts between instances (`from`, `to`).

## **2. Build weighted part‑type interaction table**

library(dplyr)

# Map instance IDs to their part types

inst_to_part <- world_parts %>% select(instance_id, part)

edges_part \<- edges_instances %\>% left_join(inst_to_part, by = c("from" =
"instance_id")) %\>% rename(part_from = part) %\>% left_join(inst_to_part, by = c("to" =
"instance_id")) %\>% rename(part_to = part) %\>% \# drop self-contacts between instances
of same type if you want only heterotypic \# filter(part_from != part_to) %\>% mutate( \#
undirected canonical ordering of part-pair p1 = pmin(part_from, part_to), p2 =
pmax(part_from, part_to) ) %\>% group_by(p1, p2) %\>% summarise( weight = n(), \# number
of instance-level contacts mean_distance = mean(distance), \# optional summary .groups =
"drop" )

edges_part\`

## **3. Create the part‑type graph**

```         
r
```

\`library(igraph)

g_part \<- graph_from_data_frame( edges_part, directed = FALSE )

E(g_part)$weight        <- edges_part$weight
E(g_part)$mean_distance <- edges_part$mean_distance

g_part\`

## **4. Optional: add node attributes (type, color diversity, etc.)**

```         
r
```

\`# simple flag for primitive vs submodel at part level parts_attr \<- tibble( part =
V(g_part)$name
) %>%
  mutate(
    is_submodel  = str_detect(part, "\\.ldr$"), is_primitive = str_detect(part,
"\\.dat\$") )

V(g_part)$is_submodel  <- parts_attr$is_submodel[match(V(g_part)$name, parts_attr$part)]
V(g_part)$is_primitive <- parts_attr$is_primitive[match(V(g_part)$name, parts_attr$part)]\`

Now `g_part` is a weighted interaction network between LEGO part types, derived from the
underlying brick‑instance contacts.


