---
title: "test_lego_model_network"
author: "Pedro Jordano"
date: "2026-03-04"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(here)
library(stringr)
library(tidyr)
library(igraph)
library(ggnetwork); library(intergraph)
library(igraph)
library(ggraph)
library(ggplot2)
library(GGally)

library(jsonlite)  # For API
library(purrr)
library(httr)
library(ggimage)
library(grid)
library(colouR)
library("png")                      # Load png package
# install.packages("patchwork")       # Install patchwork package
library("patchwork")                # Load patchwork
library(tidygraph)

theme_set(theme_minimal())

```

## Introduction

Here is a minimal R workflow that parses this MPD and builds a simple **part–part**
contact network by treating parts that appear in the same submodel as connected. We can
refine the contact criterion later (e.g. geometric distance) if needed.

## **1. Read and parse the MPD**

Parse an `.ldr` or and `.mpd` file to get the part list.

```{r parse, echo=TRUE}
# 1) Read a .mpd file for the selected LEGO set.
set_id<- "4489-1" ###### <<<<<------- Input set code here.
set_num<- set_id
# mpd_lines <- readLines(here::here("data/models/601-2-Tow_Truck.mpd"))
# For Star Wars models
mpd_lines <- readLines(here::here(paste0("data/models/starwars/",set_id,".mpd")))

# Helper: extract setname.
source(here::here("functions/extract_mpd_set_name.R"))
# setname<- extract_mpd_set_name(here::here("data/models/starwars/7131-2-PJ.mpd"))
setname<- extract_mpd_set_name(here::here(paste0("data/models/starwars/",set_id,".mpd")))

# readPNG function to load an image of the set that we have stored on our computer:
# For later use as inset in a network graph.
my_image <- readPNG(here::here(paste0("images/starwars_sets_images/",setname,".png"))) #, 
#  native = TRUE))

# set_image_row <- sets %>%
#    dplyr::filter(set_num == set_id)
# 
# set_image_url <- set_image_row$img_url

# Helper: identify FILE sections
file_idx  <- grep("^0 FILE", mpd_lines)
file_idx_end <- c(file_idx[-1] - 1, length(mpd_lines))

files_df <- tibble(
    file_start = file_idx,
    file_end   = file_idx_end,
    file_header = mpd_lines[file_idx]
) %>%
    mutate(
        file_name = str_trim(str_remove(file_header, "^0 FILE")),
        file_name = str_replace(file_name, "^\\d+\\s+-\\s+", "")  # strip leading "106 - "
    )

files_df

```

## **2. Extract part instances (LDraw type 1 lines)**

```{r parse_1, echo=TRUE}
# Function to parse "1 ..." lines inside a FILE block
parse_file_parts <- function(lines, file_name) {
    # LDraw "1" line: 1 col x y z a b c d e f g h i part.dat or subfile.ldr
    part_lines_idx <- grep("^1\\s", lines)
    if (length(part_lines_idx) == 0) return(NULL)
    
    # Split by whitespace
    tokens <- str_split(lines[part_lines_idx], "\\s+", simplify = FALSE)
    
    # Map to data frame
    parts <- lapply(tokens, function(tok) {
        # minimum length check
        if (length(tok) < 15) return(NULL)
        tibble(
            file      = file_name,
            color     = tok[2],
            x         = as.numeric(tok[3]),
            y         = as.numeric(tok[4]),
            z         = as.numeric(tok[5]),
            a         = as.numeric(tok[6]),
            b         = as.numeric(tok[7]),
            c         = as.numeric(tok[8]),
            d         = as.numeric(tok[9]),
            e         = as.numeric(tok[10]),
            f         = as.numeric(tok[11]),
            g         = as.numeric(tok[12]),
            h         = as.numeric(tok[13]),
            i         = as.numeric(tok[14]),
            part_raw  = paste(tok[15:length(tok)], collapse = " ")
        )
    })
    
    bind_rows(parts)
}

# Apply to each FILE block
parts_all <- purrr::map2_dfr(
    files_df$file_start,
    files_df$file_end,
    ~{
        block_lines <- mpd_lines[.x:.y]
        fname <- files_df$file_name[files_df$file_start == .x]
        parse_file_parts(block_lines, fname)
    }
)

parts_all

```

This gives a table with one row per placed element (including submodels like
`106 - car.ldr`, `106 - Minifig.ldr`, and `106 - wheel assembly.ldr`).

## **3. Decide what counts as a “part” and “connection”**

For a first simple network, you can:

-   Represent each **part file** (`*.dat`, `*.ldr`) as a node.

-   Add an edge between two parts if they appear together in the same FILE (submodel).

Below I create a co‑occurrence network at the level of **part_raw** (so `3004.dat` etc.
are nodes). You can easily switch to unique instance IDs if you want a multigraph.

```{r parts_as_nodes, echo=TRUE}
# Clean part name
parts_all <- parts_all %>%
    mutate(
        part = str_trim(part_raw)
    )

# Identify file type
parts_all <- parts_all %>%
    mutate(
        is_submodel  = str_detect(part, "\\.ldr$"),
        is_primitive = str_detect(part, "\\.dat$")
    )

# Optional: drop minifig-related primitives by a simple pattern or list
minifig_patterns <- c("minifig", "head", "torso", "hip", "leg", "arm", "hand")
parts_all <- parts_all %>%
    filter(is_primitive) %>%                                # keep only .dat
    filter(!str_detect(str_to_lower(part), 
                       paste(minifig_patterns, collapse = "|")))

# Co‑occurrence edges among primitives only
edges_df <- parts_all %>%
    group_by(file) %>%
    summarise(parts = list(unique(part)), .groups = "drop") %>%
    mutate(
        edges = purrr::map(parts, ~{
            p <- .x
            if (length(p) < 2) return(NULL)
            comb <- t(combn(p, 2))
            tibble(from = comb[,1], to = comb[,2])
        })
    ) %>%
    select(edges) %>%
    unnest(edges)

edges_df

```

## **4. Build igraph object and analyze**

```{r build_graph, echo=TRUE, warning=FALSE, fig.height= 12, fig.width=18}
# Graph of primitive parts only
g <- graph_from_data_frame(edges_df, directed = FALSE)

# Vertex ids for joining to Rebrickable. Vertex ids (strip .dat).
V(g)$part_id <- sub("\\.dat$", "", V(g)$name)

# -------------------------------------------------------------------
# 2. Download/cache part images and attach paths to vertices
# -------------------------------------------------------------------
target_set <- set_id
api_key    <- "40c6692458e8b4c27074486ab114d6cc"

source(here::here("functions/get_LEGO_set_parts.R"))
target_set_parts <- get_LEGO_set_parts(target_set, api_key)

source(here::here("functions/get_LEGO_part_images_for_set.R"))
target_set_parts <- get_LEGO_part_images_for_set(target_set)

# One row per part_num
target_set_parts <- target_set_parts %>%
  dplyr::distinct(part_num, .keep_all = TRUE)

# Attach local img path to vertices (by part_id)
nodes_tbl <- tibble::tibble(part_id = V(g)$part_id) %>%
  dplyr::left_join(
    target_set_parts %>%
      dplyr::select(part_num, img_local_path),
    by = c("part_id" = "part_num")
  )

# Attach local img path to vertices (by part_id)
nodes_tbl <- tibble::tibble(part_id = V(g)$part_id) %>%
  dplyr::left_join(
    target_set_parts %>%
      dplyr::select(part_num, img_local_path),
    by = c("part_id" = "part_num")
  )

# Manually fix missing images for specific parts
nodes_tbl <- nodes_tbl %>%
  mutate(
    img_local_path = dplyr::case_when(
      part_id == "4592"  ~ "images/part_images/4592.png",
      part_id == "4593"  ~ "images/part_images/4593.png",
      part_id == "44567" ~ "images/part_images/44567a.png",
      TRUE               ~ img_local_path
    )
  )

stopifnot(nrow(nodes_tbl) == igraph::vcount(g))

V(g)$img_path <- nodes_tbl$img_local_path

# After: V(g)$img_path <- nodes_tbl$img_local_path

missing_imgs <- tibble(
  part_id = V(g)$part_id,
  img_path = V(g)$img_path
) %>%
  filter(is.na(img_path))

missing_imgs

# To see just which part_ids are missing:
unique(missing_imgs$part_id)

# -------------------------------------------------------------------
# 3. Plot g with small part images as nodes
# -------------------------------------------------------------------
tg <- tidygraph::as_tbl_graph(g)
set.seed(123)
layout_df <- ggraph::create_layout(tg, layout = "fr")

gg_lego <- ggraph::ggraph(layout_df) +
      ggraph::geom_edge_bend0(width = 0.38, alpha = 0.4) +
      ggraph::geom_node_point(size = 0, alpha = 0) +
      ggimage::geom_image(
      ggplot2::aes(x = x, y = y, image = img_path),
    size = 0.05,
    inherit.aes = FALSE
  ) +
  ggraph::geom_node_text(
    ggplot2::aes(label = part_id),
#    nudge_y = -0.03,
    vjust = 4.25,
    size    = 5,
    color   = "black",
#    family  = "NULL"        # empty family = device default; safest for PDF
  ) +
  ggraph::theme_graph() +
  ggplot2::labs(title = paste0("LEGO Parts Network ", setname))

gg_lego

```

Inserting the set image.

```{r set_image, echo=TRUE, warning=FALSE, fig.height= 12, fig.width=18}
# Optional inset: keep fonts implicit (no family argument)
my_image_array <- png::readPNG(
  here::here(paste0("images/starwars_sets_images/", setname, ".png"))
)

p_logo <- ggplot2::ggplot() +
  ggplot2::annotation_custom(
    grid::rasterGrob(my_image_array, interpolate = TRUE),
    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
  ) +
  ggplot2::theme_void()

gg_lego_inset <- gg_lego +
  patchwork::inset_element(
    p_logo,
    left   = 0.95,
    bottom = 0.6,
    right  = 0.19,
    top    = 0.8,
    align_to = "full",
    clip     = TRUE
  )

gg_lego_inset

```


-----------------------------------------------------------------------------------------------
## Session Info

```{r session, echo=TRUE}
session_info()

```

