---
title: "test_lego_model_network"
author: "Pedro Jordano"
date: "2026-03-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(here)
library(stringr)
library(tidyr)
library(igraph)
library(ggnetwork); library(intergraph)
library(igraph)
library(ggraph)
library(ggplot2)
library(GGally)

library(jsonlite)  # For API
library(purrr)
library(httr)
library(ggimage)

```

## Introduction

Here is a minimal R workflow that parses this MPD and builds a simple **part–part**
contact network by treating parts that appear in the same submodel as connected. You can
refine the contact criterion later (e.g. geometric distance) if you wish.

## **1. Read and parse the MPD**

Parse an `.ldr` or and `.mpd` file to get the part list.

```{r parse, echo=TRUE}
# 1) Read file
mpd_lines <- readLines(here::here("connectivity_workflow/4489-1-AT-AT-Mini.mpd"))

# Helper: identify FILE sections
file_idx  <- grep("^0 FILE", mpd_lines)
file_idx_end <- c(file_idx[-1] - 1, length(mpd_lines))

files_df <- tibble(
    file_start = file_idx,
    file_end   = file_idx_end,
    file_header = mpd_lines[file_idx]
) %>%
    mutate(
        file_name = str_trim(str_remove(file_header, "^0 FILE")),
        file_name = str_replace(file_name, "^\\d+\\s+-\\s+", "")  # strip leading "106 - "
    )

files_df

```

## **2. Extract part instances (LDraw type 1 lines)**

You can also embed plots, for example:

```{r parse_1, echo=TRUE}
# Function to parse "1 ..." lines inside a FILE block
parse_file_parts <- function(lines, file_name) {
  # LDraw "1" line: 1 col x y z a b c d e f g h i part.dat or subfile.ldr
  part_lines_idx <- grep("^1\\s", lines)
  if (length(part_lines_idx) == 0) return(NULL)

  # Split by whitespace
  tokens <- str_split(lines[part_lines_idx], "\\s+", simplify = FALSE)

  # Map to data frame
  parts <- lapply(tokens, function(tok) {
    # minimum length check
    if (length(tok) < 15) return(NULL)
    tibble(
      file      = file_name,
      color     = tok[2],
      x         = as.numeric(tok[3]),
      y         = as.numeric(tok[4]),
      z         = as.numeric(tok[5]),
      a         = as.numeric(tok[6]),
      b         = as.numeric(tok[7]),
      c         = as.numeric(tok[8]),
      d         = as.numeric(tok[9]),
      e         = as.numeric(tok[10]),
      f         = as.numeric(tok[11]),
      g         = as.numeric(tok[12]),
      h         = as.numeric(tok[13]),
      i         = as.numeric(tok[14]),
      part_raw  = paste(tok[15:length(tok)], collapse = " ")
    )
  })

  bind_rows(parts)
}

# Apply to each FILE block
parts_all <- purrr::map2_dfr(
  files_df$file_start,
  files_df$file_end,
  ~{
    block_lines <- mpd_lines[.x:.y]
    fname <- files_df$file_name[files_df$file_start == .x]
    parse_file_parts(block_lines, fname)
  }
)

parts_all

```

This gives a table with one row per placed element (including submodels like
`106 - car.ldr`, `106 - Minifig.ldr`, and `106 - wheel assembly.ldr`).

## **3. Decide what counts as a “part” and “connection”**

For a first simple network, you can:

-   Represent each **part file** (`*.dat`, `*.ldr`) as a node.

-   Add an edge between two parts if they appear together in the same FILE (submodel).

Below I create a co‑occurrence network at the level of **part_raw** (so `3004.dat` etc.
are nodes). You can easily switch to unique instance IDs if you want a multigraph.

```{r parts_as_nodes, echo=TRUE}
# 3) Clean part name (strip leading "106 - " from submodels, keep .dat etc.)
parts_all <- parts_all %>%
  mutate(
    part = str_trim(part_raw)
  )

# 4) Build co-occurrence edges:
# For each FILE, connect all distinct part types that co-occur
edges_df <- parts_all %>%
  group_by(file) %>%
  summarise(parts = list(unique(part)), .groups = "drop") %>%
  mutate(
    edges = purrr::map(parts, ~{
      p <- .x
      if (length(p) < 2) return(NULL)
      # all unordered pairs
      comb <- t(combn(p, 2))
      tibble(from = comb[,1], to = comb[,2])
    })
  ) %>%
  select(edges) %>%
  unnest(edges)

edges_df

```

## **4. Build igraph object and analyze**

```{r build_graph, echo=TRUE}
# 5) Create graph
g <- graph_from_data_frame(edges_df, directed = FALSE)

# Basic stats
print(g)
degree(g)

# Example: plot
plot(
  g,
  vertex.size = 5,
  vertex.label.cex = 0.6,
  vertex.label.dist = 0.5,
  main = "LEGO parts co-occurrence network"
)

# Using ggraph (Recommended)
# ggraph extends ggplot2 for networks and works directly with igraph.

ggraph(g, layout = "fr") +  # Fruchterman-Reingold layout
# geom_edge_link(aes(width = 0.0005), alpha = 0.2) +
  geom_edge_link(aes(alpha= 0.6)) +
# geom_edge_bend0() +
  geom_node_point(aes(size = degree(g)), color = "blue") +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_graph() +
  labs(title = "igraph Network with ggraph")

```

## **5. Optional: distinguish submodels vs primitive parts**

We can tag nodes as “submodel” when they are `.ldr` vs `.dat`, which might help link to
assembly hierarchy.

```{r submodels, echo=TRUE}
node_attr <- tibble(
  part = V(g)$name
) %>%
  mutate(
    is_submodel = str_detect(part, "\\.ldr$"),
    is_primitive = str_detect(part, "\\.dat$")
  )

V(g)$is_submodel  <- node_attr$is_submodel[match(V(g)$name, node_attr$part)]
V(g)$is_primitive <- node_attr$is_primitive[match(V(g)$name, node_attr$part)]

```

We may prefer the network at the level of **individual brick instances** (each physical
brick a node) with edges defined by spatial proximity, or is this part‑type co‑occurrence
representation what we want?

### See file `LEGO_model_networks_ind-parts.Rmd`
